---
title: "HW 7: Stability"
author: "Hannah Irish & Nadine Snyder"
date: "2023-06-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(deSolve)
library(cowplot)

source(here("R/lotvmodK.R"))
```

### Stability


Set initial conditions
```{r}

##parameters
pars = data.frame(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000)

##over 500 days
days = seq(from=1, to=500)

##current population
currpop = c(prey=90, pred=30)

```


No Hunting:
```{r}

##with initial function and no hunting
initial = ode(func=lotvmodK, y=currpop, times=days, parms=pars)

##make data frame
initial_df = as.data.frame(initial) %>% gather(key="animal", value="pop",-time)

##make plot
initial_plot <- ggplot(initial_df, aes(time, pop, col=animal))+
  geom_line() +
  theme_minimal()
```

Vary initial populations:

```{r}

##with different populations

##even populations of prey and predator
currpop = c(prey=30, pred=30)

same_start = ode(func=lotvmodK, y=currpop, times=days, parms=pars)

same_start_df = as.data.frame(same_start) %>% gather(key="animal", value="pop",-time)

same_start_plot <- ggplot(same_start_df, aes(time, pop, col=animal))+
  geom_line() +
  theme_minimal()


##4x as many prey
currpop = c(prey=200, pred=50)

manyprey = ode(func=lotvmodK, y=currpop, times=days, parms=pars)

manyprey_df = as.data.frame(manyprey) %>% gather(key="animal", value="pop",-time)

manyprey_plot <- ggplot(manyprey_df, aes(time, pop, col=animal))+
  geom_line() +
  theme_minimal()

#5x as many predators
currpop = c(prey=200, pred=1000)

manypred = ode(func=lotvmodK, y=currpop, times=days, parms=pars)

manypred_df = as.data.frame(manypred) %>% gather(key="animal", value="pop",-time)

manypred_plot <- ggplot(manypred_df, aes(time, pop, col=animal))+
  geom_line() +
  theme_minimal()

##Plot

plot_grid(initial_plot, same_start_plot, manyprey_plot, manypred_plot)



```
Both populations stay fairly stable under the original Lotka Volterra model with the given interaction rate. We estimate a hunting rate on the lower side will help keep stability so we don't lose too many prey.

With how we added the hunting rate, if there are enough prey, X% (X= hunting rate) of the population will die from being hunted by predators.

Run ODE Solver with hunt rate of 20%
```{r}
##add hunting rate of 20%

pars = data.frame(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, huntr = 0.2)
res = ode(func=lotvmodK_hunt, y=currpop, times=days, parms=pars)

res_df = as.data.frame(res) %>% gather(key="animal", value="pop",-time)

p1=ggplot(res_df, aes(time, pop, col=animal))+
  geom_line() +
  theme_minimal()

p1
```

We will define stability as each population being maximum 50% of the other population (cannot be more than double the other population) and both populations must still have members by the end of the time period (can't die out)


Try different hunting rates
```{r}

##hunting rate of 90% then 70%, 50%, 30%, 10%
pars$huntr = 0.9

nine = ode(func=lotvmodK_hunt, y=currpop, times=days, parms=pars)

nine_df = as.data.frame(nine) %>% gather(key="animal", value="pop",-time)

nine_plot = ggplot(nine_df, aes(time, pop, col=animal))+
  geom_line() +
  labs(title = "Hunt Rate of 90%") +
  theme_minimal()

pars$huntr = 0.7

seven = ode(func=lotvmodK_hunt, y=currpop, times=days, parms=pars)

seven_df = as.data.frame(seven) %>% gather(key="animal", value="pop",-time)

seven_plot = ggplot(seven_df, aes(time, pop, col=animal))+
  geom_line() +
  labs(title = "Hunt Rate of 70%") +
  theme_minimal()

pars$huntr = 0.5

five = ode(func=lotvmodK_hunt, y=currpop, times=days, parms=pars)

five_df = as.data.frame(five) %>% gather(key="animal", value="pop",-time)

five_plot = ggplot(five_df, aes(time, pop, col=animal))+
  geom_line() +
  labs(title = "Hunt Rate of 50%") +
  theme_minimal()

pars$huntr = 0.3

three = ode(func=lotvmodK_hunt, y=currpop, times=days, parms=pars)

three_df = as.data.frame(three) %>% gather(key="animal", value="pop",-time)

three_plot = ggplot(three_df, aes(time, pop, col=animal))+
  geom_line() +
  labs(title = "Hunt Rate of 30%") +
  theme_minimal()

pars$huntr = 0.1

one = ode(func=lotvmodK_hunt, y=currpop, times=days, parms=pars)

one_df = as.data.frame(one) %>% gather(key="animal", value="pop",-time)

one_plot = ggplot(one_df, aes(time, pop, col=animal))+
  geom_line() +
  labs(title = "Hunt Rate of 10%") +
  theme_minimal()

cowplot::plot_grid(nine_plot, seven_plot, five_plot, three_plot, one_plot, cols = 2)


```
Higher hunting rates do not satisfy stability. Predators eat too many prey and pretty much die out as a result.

Once the hunting rate gets down to about 50% and below, the conditions start satisfying stability. 

For sustainable hunting, predators shouldn't be able to eliminate more than 50% of the prey population in a given time period at an interaction rate of 0.01.



