---
title: "Homework 3"
author: "Hannah Irish & Nadine Snyder"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message= FALSE)
library(tidyverse)
library(purrr)
source(here::here("R/yields.R"))
source(here::here("R/alm_prof.R"))

```


Read in Data
```{r}

##read in almond data
clim_data <- read.table(file = here::here("assignments/data/clim.txt"), header = TRUE)

```

Use almond profit function on the data given in assignment 2 with a price per almond ton of $4
```{r}

##make sure alm_prof function works by using it in a simple case
initial_profs <- alm_prof(price=4, cost=3.5,dataset=clim_data)
```

Running the function the first time, we get that the min yield profit is `r initial_profs[[1]]`, the max yield profit is `r initial_profs[[2]]` and the mean yield profit is `r initial_profs[[3]]`



### Sensitivity Analysis

We will test the profit function for some different almond selling prices and to see how much it changes the profit for min, max, mean almond yields, since the demand for almonds might shift the market price for almonds one way or another

```{r}

##make vector of different prices
price <- rnorm(mean=5, sd = 0.1, n = 20)
##make vector of different costs
cost <- rnorm(mean=3.5, sd = 0.05, n = 20)


##combined sensitivity analysis

price_cost <- cbind.data.frame(price, cost)

results <- price_cost %>%
  pmap(alm_prof,dataset=clim_data)


results[[1]]

mean_prof = map_df(results,`[`, c("mean")) 
mean_prof = cbind(mean_prof, price_cost)

min_prof = map_df(results,`[`, c("min"))
min_prof = cbind(min_prof, price_cost)

max_prof = map_df(results,`[`, c("max")) 
max_prof = cbind(max_prof, price_cost)

```

```{r}
p1 = ggplot(mean_prof, aes(price, mean, col=cost))+
geom_point(cex=2) +
  labs(x="Price", y="Profit", title="Mean Yield Anomaly")+
  theme_minimal()


p2 = ggplot(min_prof, aes(price, min, col=cost))+
geom_point(cex=2) +
  labs(x="Price", y="Profit", title="Min Yield Anomaly")+
  theme_minimal()


p3 = ggplot(max_prof, aes(price, max, col=cost))+
geom_point(cex=2) +
  labs(x="Price", y="Profit", title="Max Yield Anomaly")+
  theme_minimal()

c1 = ggplot(mean_prof, aes(cost, mean, col=price))+
geom_point(cex=2) +
  labs(x="Cost", y="Profit", title="Mean Yield Anomaly")+
  theme_minimal()


c2 = ggplot(min_prof, aes(cost, min, col=price))+
geom_point(cex=2) +
  labs(x="Cost", y="Profit", title="Min Yield Anomaly")+
  theme_minimal()


c3 = ggplot(max_prof, aes(cost, max, col=price))+
geom_point(cex=2) +
  labs(x="Cost", y="Profit", title="Max Yield Anomaly")+
  theme_minimal()

cowplot::plot_grid(p1,c2,p3,c1,p2,c3, nrow = 3)



```

```{r}
bp_p1 = ggplot(mean_prof, aes(price, mean, col=cost))+
geom_boxplot(color="cyan") +
  labs(x="Price", y="Profit", title="Mean Yield Anomaly")+
  theme_minimal()


bp_p2 = ggplot(min_prof, aes(price, min, col=cost))+
geom_boxplot(color="blue") +
  labs(x="Price", y="Profit", title="Min Yield Anomaly")+
  theme_minimal()


bp_p3 = ggplot(max_prof, aes(price, max, col=cost))+
geom_boxplot(color="purple") +
  labs(x="Price", y="Profit", title="Max Yield Anomaly")+
  theme_minimal()


bp_c1 = ggplot(mean_prof, aes(cost, mean, col=cost))+
geom_boxplot(color="cyan") +
  labs(x="Cost", y="Profit", title="Mean Yield Anomaly")+
  theme_minimal()

bp_c2 = ggplot(min_prof, aes(cost, min))+
geom_boxplot(color="blue") +
  labs(x="Cost", y="Profit", title="Min Yield Anomaly")+
  theme_minimal()

bp_c3 = ggplot(max_prof, aes(cost, max, col=cost))+
geom_boxplot(color="purple") +
  labs(x="Cost", y="Profit", title="Max Yield Anomaly")+
  theme_minimal()

cowplot::plot_grid(bp_p1,bp_c1,bp_p2,bp_c2,bp_p3,bp_c3, nrow = 3)

```

Based on our analysis, it looks like Price is the more important parameter for the Almond Profit
