---
title: 'Assignment 6: Sobel with ODE'
author: "Hannah Irish & Nadine Snyder"
date: "2023-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(deSolve)
library(sensitivity)
library(here)
source(here("R/forest_growth.R"))

```


Run first iteration with given parameter levels
```{r}

##set parameters (carrying capacity (K) of 250, growth rate of 1%, linear growth rate of 2)
parms = list(K=250, r=0.01, g=2)

##set initial size of forest (10kg/C)
init_size = 10

##choose time period of 300 years to run the model
tm = seq(from=1, to =300)

##run the model for 300 years 
forest = data.frame(ode(y=init_size, times=tm, forest_growth, parms=parms))
colnames(forest)[2] = "C"

```

Graph Results
```{r}
##Plot with threshold line 
ggplot(forest, aes(time, C))+
  geom_point() + 
  geom_hline(yintercept = 50, color ="red", label="Threshold") + 
  annotate("text", x=9, y=60, label="Threshold", color="red")+
  labs(x= "Time (Year)", y="Forest Size (kg/C)", title= "Trajectory of Forest Parameters with No Uncertainty")+
  theme_minimal()
```

```{r}

#set initial population
Cinitial=10 

# want to learn about sensitivity to growth rate (r), linear growth rate (g) and carrying capacity (K)
# set the number of parameters
np=2000
K = rnorm(mean=250, sd=0.1, n=np)
r = rnorm(mean=0.01, sd=0.1, n=np)
g= rnorm(mean=2, sd=0.1, n=np)
X1 = cbind.data.frame(r=r, K=K, g=g)

# repeat to get our second set of samples
K = rnorm(mean=200, sd=0.1, n=np)
r = rnorm(mean=0.05, sd=0.1, n=np)
g= rnorm(mean=2, sd=0.1, n=np)
X2 = cbind.data.frame(r=r, K=K, g=g)

# fix any negative values and they are not meaningful
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# create our sobel object and get sets of parameters for running the model

sens_C = sobolSalt(model = NULL,X1, X2, nboot = 300)


# add colnames 
colnames(sens_C$X) = c("r","K", "g")

# gets results for 300 years (evaluating every year)
simtimes = seq(from=1, to=300)
parms = list(r=sens_C$X[1,"r"], K=sens_C$X[1,"K"], g=sens_C$X[1, "g"])


colnames(result)=c("time","C")
# turn it into a data frame
result = as.data.frame(result)


# maximum population it gets to
maxC = max(result$C)

```



```{r sen2}
# metrics function

compute_metrics = function(result, thresh) {
  maxC = max(result$C)
idx = which(result$C > thresh)[1]
idx = ifelse(is.na(idx), length(result$C), idx)
threshyear = result$time[idx]
return(list(maxC=maxC, threshyear=threshyear))}


# wrapper function to do everything

# making threshold double the canopy closure threshold

C_wrapper = function(r,K,g, Cinitial, simtimes, func) {
    parms = list(r=r, K=K, g=g)
    result = ode(y=Cinitial, times=simtimes, func=func, parms=parms) 
    colnames(result)=c("time","C")
  # get metrics
  metrics=compute_metrics(as.data.frame(result), thresh=100)
  return(metrics)
}

# use pmap

allresults = as.data.frame(sens_C$X) %>% pmap(C_wrapper, Cinitial=Cinitial, simtimes=simtimes, func=forest_growth)

# extract out results from pmap into a data frame
allres = allresults %>% map_dfr(`[`,c("maxC","threshyear"))


# create boxplots
tmp = allres %>% pivot_longer(cols=everything(),names_to="metric", values_to="value")
ggplot(tmp, aes(metric, value, col=metric))+geom_boxplot()
```



```{r sen3}

sens_C_maxC = sensitivity::tell(sens_C,allres$maxC)

# first-order indices (main effect without co-variance)
sens_C_maxC$S

# total sensitivity index -note that this partitions the output variance 
sens_C_maxC$T



# create another one for max year
sens_C_threshyear = sensitivity::tell(sens_C,allres$threshyear)
# first-order indices (main effect without co-variance)
sens_C_threshyear$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
sens_C_threshyear$T

```



If climate change worsens, it is likely that the growth rate of the forest might be smaller, and it will reach the threshold growth in a longer time (or not reach it at all). Alternately, if K were smaller, the canopy time would be a fraction of what we found for max carrying capacity and the change to linear growth would happen at an earlier stage.



